

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
library(magrittr)
library(reshape2)
library(ggplot2)
library(scales)
library(MASS)
library(gganimate)
library(plyr)
```


## Description of data

There are 140 samples consisting T-cell DNA (monogenic or full spectrum) with or without 260 unique synthetic (spike-in) DNA sequences, controls and blood and tumor spike dilution series. Sequences are amplified in a multiplex PCR reaction using 20 forward and 13 reverse primers. Forward primers bind to the V region of T-cell CDR3 sequences, and reverse primers bind to the J region of the same sequence. The 260 combinations of forward and reverse primers act to amplify all possible CDR3 rearrangements that occur within a given sample, as well as all of the spikes.  

```{r}
#Set Directory

Clone_Counts <- "./data/fullBatch_DNA160609LC/normalized_clones"
Spike_Counts <- "./data/fullBatch_DNA160609LC/25bp"
Run5_MetaData <- "./data/run5_sample_description.txt"
start<-1
end<-139
#function to read tables
readTabs<-function(filename)
{
  d<-read.table(file.path(Clone_Counts, filename), sep='\t', header = T)
  d<-d[,c(1,2,46,47)]
  
}

readCommas<-function(filename)
{
  d<-read.table(file.path(Spike_Counts, filename), sep=',', header = T)
  d<-d[,3:5]
  d[,1] <- gsub('-', '', d[,1])
  d
}

# Read in files

all.clones <- list.files(Clone_Counts)
all.clones <- all.clones[order(as.numeric(gsub(".*_S|.alignment.*", '', all.clones)))]
CloneCounts <- lapply(all.clones[start:end], readTabs)


all.spikes <- list.files(Spike_Counts, pattern = ".txt$")
all.spikes <- all.spikes[order(as.numeric(gsub(".*_S|.assembled.*", '', all.spikes)))]
SpikeCounts <- lapply(all.spikes[start:end], readCommas)
Run5MetaData <- read.table(Run5_MetaData, header = T, sep = '\t', stringsAsFactors = F)


```

```{r}
#Defining lambda function as the scaling factor under independence; so that consructing VxJ table filled with lambda: expected counts under independence over median of the observed spike counts

lambda<-function(d)
{
 d.med<-median(d$spike.count)
 vbyj.first<- acast(d, V~J)
 chi<-chisq.test(vbyj.first, correct = F)
 d<- chi$expected/d.med
 lambda<-melt(d, varnames=c("V.segments","J.segments"), value.var = c("spike.count"), value.name = "lambda")
}

lambdas<-lapply(SpikeCounts, lambda)
```

```{r}
# Constructing tables with normalized clonotype counts by dividing original clone count ratios into corresponding lambdas

summ=data.frame(sample=integer(end),clonetypes=integer(end),spikes=integer(end))
normalize<-function(i)
{ 
  summ$sample[i]<<-i
  summ$clonetypes[i]<<-sum(CloneCounts[[i]]$Clone.count)
  summ$spikes[i]<<-sum(SpikeCounts[[i]]$spike.count)
  
  merged<-merge(CloneCounts[[i]],lambdas[[i]]) 
  merged$countratio<-merged$Clone.count/(summ$clonetypes[i]+summ$spikes[i])
  merged$normalized<-merged$countratio/merged$lambda 
  merged
} 

for(i in start:end)
{
  CloneCounts[[i]]<-normalize(i)
}
```




```{r}
# Taking the Blood Tissues and formatting column names
## For Blood Tissue take samples 91-111 ##
## For Tumor Tissue take samples 112-132 ##


cc.subselect<-function(i)
{
  sub<-CloneCounts[[i]][,c("Clone.ID","normalized","countratio")]
  sub<-rename(sub,c("normalized"=paste("normalized",i, sep="~"),"countratio"=paste("countratio",i, sep="~")))
  sub
}

cc.merge<-function(first,last)
{
merged<-cc.subselect(first) 

for(i in (first+1):last)
{
    merged<-merge(merged,cc.subselect(i) ,by="Clone.ID", all=TRUE)
}
merged[is.na(merged)]<-0  
merged  
}

mergedNormalized<-cc.merge(91,111)
# mergedNormalized[is.na(mergedNormalized)]<-0

```


```{r}
#Merging with metadata for sketching boxplots

longN<-function(merged)
{
longNorm<-melt(merged, id= c("Clone.ID"))
longNorm$Clone.ID<-as.numeric(longNorm$Clone.ID)
longNorm$type <- as.factor(gsub("~.+$", "", longNorm$variable))
longNorm$sample <- as.factor(as.numeric(gsub("^.+~", "", longNorm$variable)))
longNorm$variable <- NULL
longNorm<-merge(longNorm,Run5MetaData,by.x = "sample", by.y="BARCODE")
}
longNorm<-longN(mergedNormalized)
```

```{r}
# LinePlot of Total Counts over increasing ST concentration series 
## For Tumor Tissue take samples 112-132 ##

summdils<-merge(summ[91:111,],Run5MetaData,by.x = "sample", by.y="BARCODE")

base<-ggplot(summdils,aes(x=dil, y=(clonetypes+spikes)))+ ylim(0, 1250000) + scale_x_continuous(trans = 'log2', labels=scales::percent)+ ylab("Total Counts")+ ggtitle("Blood Tissue Total Counts in Increasing ST Concentrations") +geom_point(shape=1)+stat_summary(fun.y=median, geom="line", size=1)
base

base<-ggplot(summdils,aes(x=dil, y=clonetypes))+ ylim(0, 1250000) + scale_x_continuous(trans = 'log2', labels=scales::percent)+ ylab("Clonotype Counts")+ ggtitle("Blood Tissue Total Clonotype Counts in Increasing ST Concentrations") +geom_point(shape=1)+stat_summary(fun.y=median, geom="line", size=1)
base

base<-ggplot(summdils,aes(x=dil, y=spikes))+ ylim(0, 1250000) + scale_x_continuous(trans = 'log2', labels=scales::percent)+ ylab("ST Counts")+ ggtitle("Blood Tissue Total ST Counts in Increasing ST Concentrations") +geom_point(shape=1)+stat_summary(fun.y=median, geom="line", size=1)
base

base<-ggplot(summdils,aes(x=dil, y=clonetypes/(clonetypes+spikes)))+ scale_y_continuous(trans = 'log10', labels=trans_format('log10',math_format(10^.x)))+ scale_x_continuous(trans = 'log2', labels=scales::percent)+ ylab("Log Clonotype Count Ratios")+ ggtitle("Blood Tissue Total Clonotype Count Ratios in Increasing ST Concentrations") +geom_point(shape=1)+stat_summary(fun.y=median, geom="line", size=1)
base

base<-ggplot(summdils,aes(x=dil, y=spikes/(clonetypes+spikes)))+ scale_y_continuous(trans = 'log10', labels=trans_format('log10',math_format(10^.x)))+ scale_x_continuous(trans = 'log2', labels=scales::percent)+ ylab("Log ST Count Ratios")+ ggtitle("Blood Tissue Total ST Count Ratios in Increasing ST Concentrations") +geom_point(shape=1)+stat_summary(fun.y=median, geom="line", size=1)
base

```
